Modelling BCG vaccination in the UK

### 2019-12-02 rmds/01-introduction.Rmd
introduction corrections

Mechanistic mathematical models provide an assumption based framework for understanding the transmission of infectious diseases.[@Anderson1991] Mechanistic models can be used to simplify complex real-world systems, whilst retaining a linkage to real-world processes.[@Keeling2007] This is unlike statistical models, which instead focus on modelling the underlying structure of the data generally without reference to the real-world processes. There are multiple mechanistic modelling approaches the most common being compartmental based models and individual based models.[@Anderson1991; @Keeling2007] Both of these approaches can be represented as a deterministic, semi-stochastic, or fully stochastic - both of the latter include a random component whilst the former does not. Recently, mechanistic models have been combined with statistical models to account for the fact that events may be only partially observed.[@King] Mechanistic models have an advantage over statistical models in that they can more easily be used to compare alternative scenarios over long periods of time for which observed data does not exist.

@@

* **Chapter \@ref(background):** Background information is given on TB and the BCG vaccine. This information helps motivate future chapters and is useful for non-subject area experts.

@@

* Abbott S, Christensen H, Brooks-Pollock E. *Reassessing the evidence for universal school-age Bacillus Calmette Guerin (BCG) vaccination in England and Wales*, doi: https://doi.org/10.1101/567511

* Abbott S, Christensen H, Welton NJ, Brooks-Pollock E. *Estimating the effect of the 2005 change in BCG policy in England: A retrospective cohort study*, doi: https://doi.org/10.1101/567511

@@

* **`prettypublisher`**: An R package that improves the R based reproducible research workflow. It provides tooling that includes: paper and figure referencing; effect size reporting; percentage reporting; P value reporting; and produces a table ready for further word processing. Used throughout this thesis. Available from GitHub using `devtools::install_github("seabbs/prettypublisher")`. Link: https://www.samabbott.co.uk/prettypublisher/

@@

* **Explore global Tuberculosis**: Developed to showcase `geTBinR` (https://www.samabbott.co.uk/getTBinR/) package functionality. This dashboard allows the interactive exploration of WHO TB data. It can also be used to generate a static, country level, report on TB epidemiology. Link: http://seabbs.co.uk/shiny/ExploreGlobalTB/

* **Explore Tuberculosis in England and Wales**: Developed to allow public Public Health England TB Notification data to be explored interactively. Key interventions are highlighted and link to trends in TB notifications. This app is used in its static form in Chapter \@ref(background). Link: http://seabbs.co.uk/shiny/TB_England_Wales/

* **Explore infectious disease models**: Developed to be used within a modelling short course at the University of Bristol (https://github.com/bristolmathmodellers/biddmodellingcourse). This dashboard allows the user to simulate and compare a variety of compartmental infectious disease models. All model code is surfaced in an easily view-able format to allow for users to develop their own models. Link: http://seabbs.co.uk/shiny/exploreidmodels/

* **Introduction to Tuberculosis models**: Developed to allow simple TB models to be explored in an interactive session. Inspired by practicals from the Introduction to TB, run by TB MAC (http://tb-mac.org/) at the 2017 Union conference. Link: http://seabbs.co.uk/shiny/intro_to_tb_models/

* **The pebble game**: Developed as a learning aid to help a general audience understand the impact of vaccination on infectious disease dynamics. Used at Green Man 2016 as part of a week of outreach work and subsequently developed further. Link: http://seabbs.co.uk/shiny/thepebblegame/

>>>>

Mechanistic mathematical models provide an assumption based framework for understanding the transmission of infectious diseases.[@Anderson1991] Mechanistic models can be used to simplify complex real-world systems, whilst retaining a linkage to real-world processes.[@Keeling2007] They are unlike statistical models, which instead focus on modelling the underlying structure of the data generally without reference to the real-world processes. There are multiple mechanistic modelling approaches, the most common being compartmental based models and individual based models.[@Anderson1991; @Keeling2007] Both of these approaches can be represented as  deterministic, semi-stochastic, or fully stochastic - both of the latter include a random component whilst the former does not. Recently, mechanistic models have been combined with statistical models to account for the fact that events may be only partially observed.[@King] Mechanistic models have an advantage over statistical models in that they can more easily be used to compare alternative scenarios over long periods of time for which observed data does not exist.

@@

* **Chapter \@ref(background):** Background information is given on TB and the BCG vaccine. This information helps motivate future chapters and may be useful for non-subject area experts.

@@

* Abbott S., Christensen H., Brooks-Pollock E. *Reassessing the evidence for universal school-age Bacillus Calmette Guerin (BCG) vaccination in England and Wales*, doi: https://doi.org/10.1101/567511

* Abbott S., Christensen H., Welton N.J., Brooks-Pollock E.. *Estimating the effect of the 2005 change in BCG policy in England: A retrospective cohort study*, doi: https://doi.org/10.1101/567511

@@

* **`prettypublisher`**: An R package that improves the R based reproducible research workflow. It provides tooling that includes: paper and figure referencing; effect size reporting; percentage reporting; p-value reporting; and produces a table ready for further word processing. Used throughout this thesis. Available from GitHub using `devtools::install_github("seabbs/prettypublisher")`. Link: https://www.samabbott.co.uk/prettypublisher/

@@

* **Explore global Tuberculosis**: Developed to showcase `geTBinR` (https://www.samabbott.co.uk/getTBinR/) package functionality. This dashboard allows the interactive exploration of WHO TB data. It can also be used to generate a static, country level, report on TB epidemiology. Link: https://github.com/seabbs/getTBinR/

* **Explore Tuberculosis in England and Wales**: Developed to allow public Public Health England TB Notification data to be explored interactively. Key interventions are highlighted and link to trends in TB notifications. This app is used in its static form in Chapter \@ref(background). Link: https://github.com/seabbs/TB_England_Wales/

* **Explore infectious disease models**: Developed to be used within a modelling short course at the University of Bristol (https://github.com/bristolmathmodellers/biddmodellingcourse). This dashboard allows the user to simulate and compare a variety of compartmental infectious disease models. All model code is surfaced in an easily view-able format to allow for users to develop their own models. Link: https://github.com/seabbs/exploreidmodels/

* **Introduction to Tuberculosis models**: Developed to allow simple TB models to be explored in an interactive session. Inspired by practicals from the Introduction to TB, run by TB MAC (http://tb-mac.org/) at the 2017 Union conference. Link: https://github.com/seabbs/intro_to_tb_models/

* **The pebble game**: Developed as a learning aid to help a general audience understand the impact of vaccination on infectious disease dynamics. Used at Green Man 2016 as part of a week of outreach work and subsequently developed further. Link: https://github.com/seabbs/thepebblegame/
